---
title: "Multiome Workflow"
author: "Gary Schweickart"
date: "2024-05-22"
output: html_document
---

# Setup: Load Libraries and multiome functions
```{r setup, include=FALSE}
# Load packages
library(GenomeInfoDb)
library(Matrix)
library(Seurat)
library(Signac)
library(patchwork)
library(tidyverse)
library(dplyr)
library(readr)
library(magrittr)
library(readxl)
library(EnsDb.Hsapiens.v86)
library(BSgenome.Hsapiens.UCSC.hg38)
library(JASPAR2020) 
library(TFBSTools)
```

```{r set working dir}
set.seed(6)
```

```{r load multiome functions}
# Loud multiome functions
source("code/multiomeFunctions.R")
```


```{r mkdirs}
dir.create("output/objects")
dir.create("output/objects/lists")
dir.create("output/objects/individual")
```

# Create Seurat Object

```{r get annotations}
# get annotation 
annotations <- GetGRangesFromEnsDb(ensdb = EnsDb.Hsapiens.v86)
seqlevelsStyle(annotations) <- 'UCSC'
genome(annotations) <- "hg38"
```

```{r establish paths}
# set file paths
input.path <- "data/raw/matrix_files"
out.base <- "output/objects"
# get list of sample names
sample.names <- list.dirs(input.path, full.names = FALSE, recursive = FALSE)
```

```{r create sobjs}
# Run create_sobj function
sobj.list <- lapply(sample.names, create_sobj, input.path, out.base=paste0(out.base,"/objects"), annotations)
names(sobj.list) <- sample.names
saveRDS(sobj.list, file = paste0(out.base, "/lists/sobj.list.raw.rds"))
print("All saved.")
```

```{r metadata}
# Get sample metadata
md <- read_excel(paste0(out.base,"/DAPH_SampleTracking_Trunc.xlsx"))
md$race[md$race == "Black or African-American"] <- "African-American"
rownames(md) <- md$sample

for (name in names(sobj.list)) {
  sobj.list[[name]]@meta.data$Age_years <- md[name, 'AGEYEARS'][[1]]
  sobj.list[[name]]@meta.data$Age_days <- md[name, 'AGEDAYS'][[1]]
  sobj.list[[name]]@meta.data$Sex <- md[name, 'sex'][[1]]
  sobj.list[[name]]@meta.data$Race <- md[name, 'race'][[1]]
  sobj.list[[name]]@meta.data$sample_id <- name
  age <- sobj.list[[name]]@meta.data$Age_years[1]
  sobj.list[[name]]@meta.data$age.name <- case_when(age >= 0 & age <= 2 ~ "Infant",
                                                    age >= 4 & age <= 8 ~ "Child", 
                                                    age >= 13 & age <= 14 ~ "Adolscent")
    
}


```

# Cell Cycle Scoring

```{r ccs}
sobj.list <- lapply(sobj.list, cellCycleScoring)
names(sobj.list) <- sample.names
saveRDS(sobj.list, paste0(out.base, "/objects/lists/sobj.list.ccs.rds"))
```

# Multiome Quality Control (QC) metrics and plots
  
## QC metrics
  
```{r QC metrics}
# create output directory
dir.create(paste0(out.base, "/QC"))
# run qc function
sobj.list <- lapply(sobj.list, plotQC, frag.base.path = input.path, frag.full.path = NULL, out =paste0(out.base, "/QC"), 
                  sample=NULL, sample.column = "sample_id", find.markers= TRUE, 
                  run.macs2 = FALSE, macs2.path = NULL, group.by.macs2 = "seurat_clusters")
```
## MACS2

```{r}
dir.create(paste0(out.base, "/MACS2"))
# insert macs2 path
macs2.path <- ""
sobj.list <-lapply(sobj.list, runMACS2, macs2.path= macs2.path, frag.base.path = input.path,
                   frag.full.path = NULL, out=paste0(out.base, "/MACS2"), 
                   group.by="seurat_clusters", sample.column = "sample_id",
                   find.markers=TRUE, annotations=annotations)
```

```{r}
# save list
saveRDS(sobj.list, paste0(out.base, "/objects/lists/sobj.list.metrics.rds"))
```

## Merge Seurat Objects

```{r merge}
dir.create(paste0(out.base, "/objects/merged"))
sobj.merged <- merge(sobj.list[[1]], sobj.list[2:length(sobj.list)])
sobj.merged <- JoinLayers(sobj.merged)
saveRDS(sobj.merged, paste0(out.base,"/objects/merged/sobj.merged.rds"))
```

## Merged QC

```{r}
dir.create(paste0(out.base, "/QC/merged"))
sobj.merged <- plotQC(sobj = sobj.merged, out = paste0(out.base, "/QC/merged"), merged = TRUE, sample = NULL, sample.column = NULL)
```


## Scrublet

First, we create a directory full of the features (genes) tab separated value files.
```{bash}
# make genes folder
mkdir output/objects/pediatric_multi/genes

for d in data/raw/SC0*/; 
  do 
    name=$(basename $d)
    gunzip -c data/raw/"$name"/outs/filtered_feature_bc_matrix/features.tsv.gz  >  output/objects/pediatric_multi/genes/"$name"_features.tsv
  done
```


Then we run scrublet in python (code/scrub.py).

Back to R to add doublet score to metadata.
```{r}
# Get doublet scores
sobj.merged@meta.data$doublet_score <- "0"
for (sample in sample.names) {
  pscrub <- read_table(paste0(out.base,"/scrub/", sample, "_doublet_scores.txt"),col_names = F)
  
  sobj.merged@meta.data[sobj.merged@meta.data$sample_id == sample, "doublet_score"] <- pscrub$X1
}
sobj.merged@meta.data$Predicted_Doublets <- ifelse(sobj.merged$doublet_score > 0.25, "Doublet","Singlet" )
table(sobj.merged@meta.data$Predicted_Doublets)

saveRDS(sobj.merged, paste0(out.base, "/objects/merged/sobj.merged.metrics.rds"))
```

## Assign QC cutoffs and subset

```{r}
sobj.merged <- subset(sobj.merged,
               subset = 
                 TSS.enrichment > 1.5 &
                 nucleosome_signal < 2.5 &
                 blacklist_ratio < 0.05 &
                 nFeature_ATAC < 30000 &
                 nCount_ATAC < 80000 &
                 nFeature_RNA < 5000 &
                 nCount_RNA < 30000 &
                 percent.mt < 20 &
                 Predicted_Doublets == "Singlet")
```


# Integrate using Harmony
```{r}
# Perform log-normalization and feature selection, as well as SCT normalization on global object
DefaultAssay(sobj.merged) <- "RNA"
sobj.merged <- sobj.merged %>%
  NormalizeData() %>%
  FindVariableFeatures(selection.method = "vst", nfeatures=3000) %>%
  ScaleData() %>%
  SCTransform(vars.to.regress = NULL)
saveRDS(sobj.merged, file = paste0(out.base, "/objects/merged/sobj.mergedSCT.RDS"))
```

### Similarly run for adult multiome and fetal multiome


